---
title: "FQA and iNaturalist Example Workflow"
author: "Iris Foxfoot"
date: "2023-06-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#load packages
library(fqacalc)
library(rinat)
library(dplyr)

#if you don't already have packages loaded, you will need to install them
#install.packages("fqacalc")

#note: fqacalc works as a stand alone package.
```

# Rstudio Basics

-   create a cell with green button in top right, code -> insert chunk, or option + CMD + i/alt + CTRL + i\
-   run a cell with green arrow or CMD + Return/Shift + Enter

# Exploring the `fqacalc` Package

```{r}
#view R package vignette (AKA tutorial)
vignette(topic = "introduction", package = "fqacalc")

#go to packages -> fqacalc in R studio

#get help for a specific function
?mean_c

#google
```

# Exploring `fqacalc` Data

```{r}
#to view what regional FQA databases are available in the package
fqa_names <- db_names()

#to view any particular list
texas <- view_db("texas_blackland_prairie_2019")
```

# Step 1. Get Observations From iNaturalist

```{r}
#get inat observations from a project using rinat package
plants <- rinat::get_inat_obs_project("blackland-prairie-nature-preserve", 
                                      type = "observations")

#note: if you were uploading your own data it might look like this
#my_data <- read.csv("path/to/data.csv")
```

# Step 2. Data Cleaning

`fqacalc` requires data inputted to be in a specific format:

1.  The data must have either a column named name containing scientific names of plant species, or a column named acronym containing acronyms of plant species.

2.  If the user is calculating cover-weighted metrics, the data must have another column containing cover values and it must be called cover.

3.  If the user is calculating cover-weighted metrics for a transect containing multiple plots, the data should also have a column containing the plot ID. The plot ID column can have any name, and it can contain numbers or characters, as long as the IDs are exactly the same within plots but distinct between plots.

| name    | cover | plot_ids |
|---------|-------|----------|
| plant A | 40    | 1        |
| plant B | 50    | 1        |
| plant A | 20    | 2        |
| plant B | 30    | 2        |
| plant C | 10    | 2        |

: Example of Data Formatting Requirements

```{r}
#do a little data cleaning
plants_clean <- plants %>% #pipelines! CMD/CTRL + Shift + m
  rename(name = taxon.name) %>% #rename column containing sci names
  filter(quality_grade == "research") %>%  #filter for research quality
  filter(iconic_taxon.name == "Plantae") %>% #filter for plants
  distinct(name) #get unique plants
```

# Step 3. Calculate FQA Metrics!

```{r}
#check names of regional fqa dbs we could use by viewing names

#see which plants match up with those in the texas FQA database
accepted <- fqacalc::accepted_entries(plants_clean, 
                                      key = "name", 
                                      db = "texas_blackland_prairie_2019", 
                                      native = F)

#calculate some metrics!
fqacalc::all_metrics(accepted, 
                     key = "name", 
                     db = "texas_blackland_prairie_2019")

#note: % hydrophytes should be NA, not 0. We are planning on fixing this!!
```
